---
title: "Model fitting"
author: "Zhang and Feng"
date: "Updated at `r lubridate::now()`"
output:
  html_notebook:
    number_section: yes
    theme: readable
    toc: yes
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
---

```{r load packages and data and basic configurations, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
# load packages
library(tidyverse)
library(lavaan)
library(semPlot)
# configurations
cfa_base <- "cfa"
data_dir <- file.path(cfa_base, "data")
mdl_spec_dir <- file.path(cfa_base, "modelspec")
res_dir <- file.path(cfa_base, "result")
# load data
ef_data <- read_csv(file.path(data_dir, "ef_behav_all_trans_regress_scaled.csv"))
# load model specifications
mdl_three <- read_lines(file.path(mdl_spec_dir, "full_three_factor.lav"))
mdl_one <- read_lines(file.path(mdl_spec_dir, "one_factor.lav"))
mdl_shft_inhb <- read_lines(file.path(mdl_spec_dir, "ShftInhb.lav"))
mdl_shft_updt <- read_lines(file.path(mdl_spec_dir, "ShftUpdt.lav"))
mdl_updt_inhb <- read_lines(file.path(mdl_spec_dir, "UpdtInhb.lav"))
mdl_bifactor <- read_lines(file.path(mdl_spec_dir, "bifactor.lav"))
mdl_bifactor_updt_shft <- read_lines(file.path(mdl_spec_dir, "bifactor_updt_shft.lav"))
mdl_bifactor_updt_inhb <- read_lines(file.path(mdl_spec_dir, "bifactor_updt_inhb.lav"))
mdl_bifactor_inhb_shft <- read_lines(file.path(mdl_spec_dir, "bifactor_inhb_shft.lav"))
mdl_bifactor_inhb <- read_lines(file.path(mdl_spec_dir, "bifactor_inhb.lav"))
mdl_bifactor_shft <- read_lines(file.path(mdl_spec_dir, "bifactor_shft.lav"))
mdl_bifactor_updt <- read_lines(file.path(mdl_spec_dir, "bifactor_updt.lav"))
# path graph plot
cfa_plot <- function(mdl){
  semPaths(mdl, whatLabels = "par", intercepts = FALSE,rotation = 4, sizeMan = 10, sizeMan2 = 6.18, sizeLat = 10, mar = c(3, 20, 3, 20), nCharNodes = 0)
}
```
# Confirmatory Factor Analysis

## Three factor model

### Full three factor model

This model assumes that there exist three correlated latent variables. More specifically, the model is:

$$Inhibition =~ AntiSac + StopSignal + Stroop$$
$$Shifting =~ CateSwitch + ShiftColor + ShiftNumber$$
$$Updating =~ KeepTrack + spatialWM + WM3 $$

#### Maximum Likelihood Estimation

```{r full three factor}
fit.three.full.ml <- cfa(mdl_three, ef_data, std.lv = TRUE, missing = "ML")
summary(fit.three.full.ml, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.three.full.ml)
```
### Independent three factor model

This model assumes that there exist three independent latent variables.

#### Maximum Likelihood Estimation

```{r independent three factor}
fit.three.indep.ml <- cfa(mdl_three, ef_data, std.lv = TRUE, missing = "ML",orthogonal = TRUE)
summary(fit.three.indep.ml, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.three.indep.ml)
```
## One factor model

This model assumes that there exists only one latent ability.

### Maximum Likelihood Estimation

```{r one factor}
fit.one.ml <- cfa(mdl_one, ef_data, std.lv = TRUE,missing = "ML")
summary(fit.one.ml, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.one.ml)
```
## Two factor model

### Shifting = Inhibition

This model assumes that there exists two latent abilities. In other words, only `Updating` is the differentiated ability.

#### Maximum Likelihood Estimation

```{r two-factor shft.inhb}
fit.shft.inhb.ml <- cfa(mdl_shft_inhb, ef_data, std.lv = TRUE, missing = "ML")
summary(fit.shft.inhb.ml, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.shft.inhb.ml)
```
### Shifting = Updating

This model assumes that there exists two latent abilities. In other words, only `Inhibition` is the differentiated ability.

#### Maximum Likelihood Estimation

```{r two-factor shft.updt}
fit.shft.updt.ml <- cfa(mdl_shft_updt, ef_data, std.lv = TRUE, missing = "ML")
summary(fit.shft.updt.ml, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.shft.updt.ml)
```
### Updating = Inhibition

This model assumes that there exists two latent abilities. In other words, only `Shifting` is the differentiated ability.

#### Maximum Likelihood Estimation

```{r two-factor updt.inhb}
fit.updt.inhb.ml <- cfa(mdl_updt_inhb, ef_data, std.lv = TRUE, missing = "ML")
summary(fit.updt.inhb.ml, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.updt.inhb.ml)
```
# Bifactor Model

## Three factors

### Maximum Likelihood Estimation

```{r bifactor model ML}
fit.bifac.ml <- cfa(mdl_bifactor, ef_data, std.lv = TRUE, missing = "ML", orthogonal = TRUE, estimator = "ML")
summary(fit.bifac.ml, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.bifac.ml)
```
## Two factors

### Updating + Inhibition

#### Maximum Likelihood Estimation

```{r bifactor model UI ML}
fit.bifac.ui.ml <- cfa(mdl_bifactor_updt_inhb, ef_data, std.lv = TRUE, missing = "ML",orthogonal = TRUE, estimator = "ML")
summary(fit.bifac.ui.ml, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.bifac.ui.ml)
```
### Updating + Shifting

#### Maximum Likelihood Estimation

```{r bifactor model US ML}
fit.bifac.us.ml <- cfa(mdl_bifactor_updt_shft, ef_data, std.lv = TRUE, missing = "ML", orthogonal = TRUE, estimator = "ML")
summary(fit.bifac.us.ml, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.bifac.us.ml)
```
### Shifting + Inhibition

#### Maximum Likelihood Estimation

```{r bifactor model IS ML}
fit.bifac.is.ml <- cfa(mdl_bifactor_inhb_shft, ef_data, std.lv = TRUE, missing = "ML", orthogonal = TRUE, estimator = "ML")
summary(fit.bifac.is.ml, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.bifac.is.ml)
```
## one factors
### inhibition
### Maximum Likelihood Estimation

```{r bifactor model inhb ML}
fit.bifac.inhb.ml <- cfa(mdl_bifactor_inhb, ef_data, std.lv = TRUE, missing = "ML", orthogonal = TRUE, estimator = "ML")
summary(fit.bifac.inhb.ml, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.bifac.inhb.ml)
```
### shifting
### Maximum Likelihood Estimation

```{r bifactor model shft ML}
fit.bifac.shft.ml <- cfa(mdl_bifactor_shft, ef_data, std.lv = TRUE, missing = "ML", orthogonal = TRUE, estimator = "ML")
summary(fit.bifac.shft.ml, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.bifac.shft.ml)
```
### updating
### Maximum Likelihood Estimation

```{r bifactor model updt ML}
fit.bifac.updt.ml <- cfa(mdl_bifactor_updt, ef_data, std.lv = TRUE, missing = "ML", orthogonal = TRUE, estimator = "ML")
summary(fit.bifac.updt.ml, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.bifac.updt.ml)
```

### Model comparison

#From the above fitting results, we know that there are three models reaching the acceptable fitting criterion. We will using **likelihood ratio test** to check which is the better model.

```{r mdoel checking bifactor}
anova(fit.three.full.ml, fit.updt.inhb.ml)
anova(fit.bifac.is.ml, fit.bifac.shft.ml)
anova(fit.bifac.us.ml, fit.bifac.shft.ml)
```
# Extract scores

#We extract the scores of each subject on each ability to correlate to brain data.

```{r extract scores}
scores.ml <- lavPredict(fit.three.full.ml, method="regression")
write_csv(as.data.frame(scores.ml),file.path(res_dir,"EFfactorScore_fullthree.csv"))
```
